---
title: "Visualize Data"
author: "Samreen Aulakh"
format: gfm
execute: 
  echo: true
  eval: true
---


<!-- If you get the error "Error in list2(na.rm = na.rm, ...) : object 'ffi_list2' not found" do the following in the Console run:  

    remove.packages("rlang")

Then Restart R to be safe. Then run:

    install.packages("rlang")

and again restart R to be safe. That should fix the problem (which is a package dependency issue).   -->



Try your code again
```{r}
#| label: setup-install
#| include: false

install.packages("rlang")

```



## Your Turn 0

Add a setup chunk that loads the tidyverse packages.

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(tidyverse)
```


```{r}
#| label: render-6-rows
head(mpg)
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
#| label: ggplot-mpg
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y= hwy))
```


## Your Turn 2

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
#| label: boxplot
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))
```


## Your Turn 3

Make a histogram of the `hwy` variable from `mpg`. Hint: do not supply a y variable.

```{r}
#| label: histogram
ggplot(data = mpg) +
  geom_histogram(mapping= aes(x = hwy))
```

## Your Turn 4

Use the help page for `geom_histogram` to make the bins 2 units wide.

```{r}
#| label: histogram-binwidth-2

ggplot(data = mpg) +
  geom_histogram(mapping= aes(x = hwy), binwidth=2,
                 color="black")
```


## Your Turn 5

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
#| label: scatterplot-color-by-class
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, 
                           y = hwy,
                           color = class),
             size = 2,
             alpha = 0.5,
             shape = "triangle") +
  theme_minimal()

```

## Help Me

What do `facet_grid()` and `facet_wrap()` do? 
(run the code, interpret, convince your group)

```{r}
#| label: add-label-help-me
# Makes a plot that the commands below will modify
q <- ggplot(mpg) + geom_point(aes(x = displ, y = hwy))

#make one row, four columns by cyl
q + facet_grid(. ~ cyl)

#makes three rows, one column by drv
q + facet_grid(drv ~ .)

#makes three rows by drv, four columns by cyl
q + facet_grid(drv ~ cyl)

#makes a graph for each value in class
q + facet_wrap(~ class)
```


## Your Turn 6

Make a bar chart `class` colored by `class`. Use the help page for `geom_bar` to choose a "color" aesthetic for class.

```{r}
#| label: barplot
ggplot(data=mpg) +
  geom_bar(aes(x= class, fill= class))+
  guides(fill="none")+
  labs(x= "class of vehicles",
       y= " number of vehicles in sample") +
  theme_get()
```

## Quiz

What will this code do?

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy)) 
# ggsave("example.jpg")
```
Add two geometries to one plot.

***

# Take aways

You can use this starter code template to make thousands of graphs with **ggplot2**.

```{r}
#| label: template
#| eval: false
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```